---
title: "Updated_Hawaiian_Soil"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r, include=FALSE}
library(ggplot2)
library(randomForest)
library(dendextend)
library(DMwR)
library(caret)
library(pROC)
library(fscaret)
library(EMA)
library(ggpubr)

df = read.csv('data/Updated_Feature_Table_11_2017.csv', header = T, sep = ',', stringsAsFactors = F)
rownames(df) = df$X
df$X = NULL
df = as.data.frame(t(df))
training = df[!(is.na(df$Culture)),]
testing_set = df[!(df$Culture == 'pos' | df$Culture == 'neg'),]


## Split, convert to numeric, and factorize 
training$Culture = as.factor(as.character(training$Culture))
train_data = as.data.frame(training[,1:16])
train_data = as.data.frame(lapply(train_data, function(x) as.numeric(as.character(x))))
Culture = training[,17]
df = cbind(train_data, Culture)
```

## Correlation of Features to outcome variable (Culture Status)
```{r, echo = FALSE}
### Observed Species
col_p = data.frame(value = numeric())

for (i in 1:length(colnames(train_data))){
  q = wilcox.test(train_data[,i]~training$Culture, exact = F)
  col_p = rbind(col_p, q$p.value)
}
colnames(col_p) = "Culture_p_values"

rownames(col_p) = colnames(train_data)
signif = subset(col_p, Culture_p_values < 0.05)
knitr::kable(signif)
```

### Directionality of Significant Features

#### Maghemite Percentage By Culture Status
```{r, echo=FALSE}
p = ggplot(data = df, aes(Culture, Maghemite, color=Maghemite)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.1) + scale_colour_gradientn(colours = myPalette(2), values=seq(0, 130, length.out=130)/130)

p + stat_compare_means(comparisons = list(c("pos", "neg")), method = "wilcox.test", paired = FALSE, label = "p.signif") + stat_compare_means(label.x = 2, label.y = 19)+ theme_bw() +
  ggtitle("Maghemite Measure by Culture Status") + theme(axis.text.y = element_text(colour="grey20",size=12,angle=0,hjust=1,vjust=0,face="plain"),
                                                                      axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=.5,face="plain"),
                                                                      axis.text.x = element_text(colour="grey20",size=12,angle=0,hjust=0.5,vjust=0,face="plain"),
                                                                      axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=.5,face="plain"))
```

#### Halloysite By Culture Status
```{r, echo=FALSE}
p = ggplot(data = df, aes(Culture, Halloysite, color=Halloysite)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.1) + scale_colour_gradientn(colours = myPalette(2), values=seq(0, 130, length.out=130)/130)

p + stat_compare_means(comparisons = list(c("pos", "neg")), method = "wilcox.test", paired = FALSE, label = "p.signif") + stat_compare_means(label.x = 2, label.y = 92)+ theme_bw() +
  ggtitle("Halloysite Measure by Culture Status") + theme(axis.text.y = element_text(colour="grey20",size=12,angle=0,hjust=1,vjust=0,face="plain"),
                                                                      axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=.5,face="plain"),
                                                                      axis.text.x = element_text(colour="grey20",size=12,angle=0,hjust=0.5,vjust=0,face="plain"),
                                                                      axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=.5,face="plain"))
```

#### Kaolinite By Culture Status
```{r, echo=FALSE}
p = ggplot(data = df, aes(Culture, Kaolinite, color=Kaolinite)) + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.1) + scale_colour_gradientn(colours = myPalette(2), values=seq(0, 130, length.out=130)/130)

p + stat_compare_means(comparisons = list(c("pos", "neg")), method = "wilcox.test", paired = FALSE, label = "p.signif") + stat_compare_means(label.x = 2, label.y = 110)+ theme_bw() +
  ggtitle("Kaolinite Measure by Culture Status") + theme(axis.text.y = element_text(colour="grey20",size=12,angle=0,hjust=1,vjust=0,face="plain"),
                                                                      axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=.5,face="plain"),
                                                                      axis.text.x = element_text(colour="grey20",size=12,angle=0,hjust=0.5,vjust=0,face="plain"),
                                                                      axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=.5,face="plain"))
```


### Correlation of Features

```{r, echo=FALSE}
cor_mat = cor(train_data, method = "spearman")

library(corrplot)
cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    diag(lowCI.mat) <- diag(uppCI.mat) <- 1
    for(i in 1:(n-1)){
        for(j in (i+1):n){
            tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
            p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
            lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
            uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
        }
    }
    return(list(p.mat, lowCI.mat, uppCI.mat))
}

res1 <- cor.mtest(train_data,0.95)

corrplot(cor_mat, p.mat = res1[[1]], sig.level=0.05, type = "lower")
```

### Cross Validation Error Rate
```{r, echo = FALSE}
## Random Forest with Cross Validation
rf_cv = rfcv(trainx = train_data, trainy = Culture, cv.fold = 10)
rf_cv$error.cv
```





```{r, echo = FALSE}
library(dendextend)
# Ward Hierarchical Clustering
d <- dist(train_data, method = "euclidean") # distance matrix
fit <- hclust(d, method="complete") 
hcd = as.dendrogram(fit)
colors_to_use = as.numeric(training$Culture)
labels_colors(hcd) = colors_to_use
plot.new()
plot(hcd, cex.lab =0.5)
legend('topright', legend = c("Negative", "Positive"),lty=1, col=c("black","red"), cex = 0.75)
```


### Random Forest and Feature Importance
```{r, echo = FALSE}
library(randomForest)

## Split into Testing and Training Set
# Define train and test sets
set.seed(111)
train.percent = .8
inTrain = rownames(train_data) %in% sample(rownames(train_data), floor(train.percent*length(rownames(train_data))))
train.data = train_data[inTrain,]
test.data = train_data[!inTrain,]
testing.data = as.matrix(test.data)
## Random Forest of training data
rf = randomForest(training$Culture ~ ., data=train_data, mtry = 3, ntree=500, importance=TRUE)

## Feature Importance
features = importance(rf, type = 2)
features[order(features[,1], decreasing = TRUE),]
```



### Table of Culture Status
```{r,echo = FALSE}
test = as.data.frame(table(training$Culture))
colnames(test) = c("Culture Status", 'Frequency')
knitr::kable(test)
```


### Balance the Data
```{r, echo=FALSE}
# SMOTE the data
library(DMwR)
training2 <- SMOTE(Culture ~ ., training, perc.over = 100, perc.under=200)
table(training2$Culture)
```


```{r, echo=FALSE}
training2$Culture = as.factor(as.character(training2$Culture))

train_data2 = as.data.frame(training2[,1:16])
train_data2 = as.data.frame(lapply(train_data2, function(x) as.numeric(as.character(x))))
response2 = training2[,17]

## Random Forest with Cross Validation
rf_cv = rfcv(trainx = train_data2, trainy = response2, cv.fold = 10)

rf_cv$error.cv
```



```{r, echo=FALSE}
## Split into Testing and Training Set
# Define train and test sets
set.seed(111)
train.percent = .8
inTrain = rownames(train_data2) %in% sample(rownames(train_data2), floor(train.percent*length(rownames(train_data2))))
train.data = train_data2[inTrain,]
test.data = train_data2[!inTrain,]
testing.data = as.matrix(test.data)
## Random Forest of training data
rf = randomForest(training$Culture ~ ., data=train_data, mtry = 3, ntree=500, importance=TRUE)
## Feature Importance
features = importance(rf, type = 2)
features[order(features[,1], decreasing = TRUE),]
```



#### Unsupervised Clustering
```{r, include = FALSE}
### Unsupervised Clustering
df=as.matrix(train_data)
wss <- (nrow(df)-1)*sum(apply(df,2,var))
for (i in 1:15) wss[i] <- sum(kmeans(df, 
  	centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
```


```{r, include=FALSE}
# Ward Hierarchical Clustering
d <- dist(train_data, method = "euclidean") # distance matrix
fit <- hclust(d, method="complete") 
hcd = as.dendrogram(fit)
plot(fit)
```
